---
title: "HELPmiss (Data Verbs) Activity"
author: "Mia Iceland"
output: html_notebook
---


## Set Up:

```{r message=FALSE}
rm(list = ls())     # clean up your R environment

# load packages
library(tidyverse)  # includes lots of data verbs like `group_by()` and `summarise()`
library(mosaicData) # includes the `HELPmiss` data set 


# Load the `HELPmiss` data set into our RStudio environment
data("HELPmiss", package = "mosaicData")

```

## Helpful links: 

- Look though the DC Textbook for `tidyverse` functions. 
- Check out some of these RStudio cheat sheets:
  - <https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf> 
  - <https://www.rstudio.com/resources/cheatsheets/>
  - <https://s3.amazonaws.com/assets.datacamp.com/blog_assets/Tidyverse+Cheat+Sheet.pdf>
  



## Task 1: Data Description

*Write several sentences (or a bullet list) describing the HELP Study and the resulting `HELPmiss` data.  Your description should investigate basic data provenance (e.g., Who, What, When, Where, Why, How), explain the setting for the data, specify what each case represents in the data, and remark on inclusion/exclusion criteria.* 

- Where: Boston
- When: 2003
- Why: a trial on a clinic's ability to pair primary care with an adult's health evaluation
- How: a baseline interview with follow-up interviews every 6 months for 2 years
- Case: adults with no primary care physician
  - they were recruited from a detoxification unit
  - went through a general assessment, then an intervention and primary care
  - 470 participants
- Data Criteria:
  - adults, speak English or Spanish
  - alcohol, heroin, or cocaine being the patient's first or second drug of choice
  - lives close to the primary care facility or are homeless
  - couldn't already have a primary care, dementia, plans to leave the area, or pregnant 


## Task 2: Basic Summaries 

*Produce one or more R expressions involving `summarize()` and `HELPmiss` to address each of the following prompts.*

1. number of people (cases) in `HELPmiss` study

```{r}
HELPmiss %>%
  summarise(count= n())

```

2. total number of times in the past 6 months entered a detox program (measured at baseline) for all the people in `HELPmiss`. 

```{r}
HELPmiss %>%
  summarise(sum(e2b, na.rm=TRUE))

```


3. mean time (in days) to first use of any substance post-detox for all the people in `HELPmiss`


```{r}
HELPmiss %>%
  summarise(mean(daysanysub, na.rm=TRUE))

```



## Task 3: Group Summaries  

*Repeat task 2 above, but add code chunks to calculate result group-by-group according to each prompt (i.e. each prompt should have three statements for each of the three prompts in task 2).  Be sure to show all R code and write a sentence or two about what you observe in the results. Remember, you can add multiple statistics inside `summary`.* 


- males versus females
  - I noticed that there are more male participants in this study than females. Also, more males than females had previously entered a detox program before 6 months before this trial started. Finally, men are faster to start reusing a substance post-detox. 
- homeless or not
  - More people in this study have a home than those that are homeless. More homeless people than people with homes had previously entered a detox program before 6 months before this trial started. Lastly, homeless people are faster to start reusing a substance post-detox.
- substance
  - The most abused substance in this trial is alcohol. Alcohol abusers are the largest group of people who had previously entered a detox program before 6 months before this trial started. Finally, people who use heroin are faster to start reusing their "substance of choice" post-detox. 
- break down the homeless versus housed further, by sex
  - The greatest number of people in this study are housed males. Also, homeless males are the largest group of people who had previously entered a detox program before 6 months before this trial started. Lastly, homeless males are faster to start reusing their abused substance post-detox. 
- homeless versus housed broken down by substance
  - The greatest number of people in this study are homeless and abuse alcohol. Homeless alcohol abusers are the largest group of people who had previously entered a detox program before 6 months before this trial started. Housed heroin users are faster to start reusing heroin post-detox. 

### males versus females
```{r}

HELPmiss %>%
  group_by(sex) %>%
  summarise(count= n())

HELPmiss %>%
  group_by(sex) %>%
  summarise(sum(e2b, na.rm=TRUE))

HELPmiss %>%
  group_by(sex) %>%
  summarise(mean(daysanysub, na.rm=TRUE))

```

### homeless or not
```{r}

HELPmiss %>%
  group_by(homeless) %>%
  summarise(count= n())

HELPmiss %>%
  group_by(homeless) %>%
  summarise(sum(e2b, na.rm=TRUE))

HELPmiss %>%
  group_by(homeless) %>%
  summarise(mean(daysanysub, na.rm=TRUE))

```


### substance
```{r}

HELPmiss %>%
  group_by(substance) %>%
  summarise(count= n())

HELPmiss %>%
  group_by(substance) %>%
  summarise(sum(e2b, na.rm=TRUE))

HELPmiss %>%
  group_by(substance) %>%
  summarise(mean(daysanysub, na.rm=TRUE))

```

### homeless versus housed broken down by sex
```{r}

HELPmiss %>%
  group_by(sex, homeless) %>%
  summarise(count= n())

HELPmiss %>%
  group_by(sex, homeless) %>%
  summarise(sum(e2b, na.rm=TRUE))

HELPmiss %>%
  group_by(sex, homeless) %>%
  summarise(mean(daysanysub, na.rm=TRUE))

```

### homeless versus housed broken down by substance
```{r}

HELPmiss %>%
  group_by(homeless, substance) %>%
  summarise(count= n())

HELPmiss %>%
  group_by(homeless, substance) %>%
  summarise(sum(e2b, na.rm=TRUE))

HELPmiss %>%
  group_by(homeless, substance) %>%
  summarise(mean(daysanysub, na.rm=TRUE))

```



## Task 4: Data Visualization & Observations

*Include one or more interesting plots from this data set involving at least 3 variables per plot. Write a few sentences to explain the story that your plot tells about these data.  You can expand on the relationships that you studied in Task 2, or you can explore a different group of variables in `HELPmiss` that show something interesting. Remember to use the interactive commands in the console, generate the R commands that will reproduce your plot, and then paste the R commands into an R chunk in the RMarkdown file.*  

*Remember, you can use* `esquisser` *or* `mplot` *in your console. But only include the ggplot code in this Rmd document.*


```{r}
# Your Code here.

```



